---
title: "Abraxas Demand Forecasting (Data Prep)"
author: "Enrique López <enrique_lcbjj@hotmail.com>"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("../")
```


## Data gathering
```{r}
require("tidyverse")
require("magrittr")
require("ggplot2")

clients <- read_csv(
  "../data/cliente_tabla.csv", 
  col_types = cols(
    Cliente_ID = col_integer()
  ))

products <- read_csv(
  "../data/producto_tabla.csv",
  col_types = cols(
    Producto_ID = col_integer()
  ))

depots <- read_csv(
  "../data/town_state_small.csv",
  col_types = cols(
    Agencia_ID = col_integer()
  )) 

train <- read_csv(
  "../data/df_[candidate]_small.csv",
  col_types = cols(
    Agencia_ID = col_integer(),
    Canal_ID = col_integer(),
    Cliente_ID = col_integer(),
    Producto_ID = col_integer()
  )) %>% mutate(ID = 1:nrow(.))

test <- read_csv(
  "../data/df_[test]_small.csv",
  col_types = cols(
    Agencia_ID = col_integer(),
    Canal_ID = col_integer(),
    Cliente_ID = col_integer(),
    Producto_ID = col_integer()
  )) %>% mutate(ID = 1:nrow(.))
```

## Data preparation

### Exploratory Analysis

```{r}
require("summarytools")
require("visdat")
```

```{r}
dfSummary(clients)
```

```{r}
dfSummary(products, 
  plain.ascii = FALSE, 
  style = "grid", 
  graph.magnif = 0.75,
  valid.col = FALSE,
  tmp.img.dir = "/tmp")
```

```{r}
dfSummary(depots,
  plain.ascii = FALSE, 
  style = "grid", 
  graph.magnif = 0.75,
  valid.col = FALSE,
  tmp.img.dir = "/tmp")
```

```{r}
dfSummary(train,
  plain.ascii = FALSE, 
  style = "grid", 
  graph.magnif = 0.75,
  valid.col = FALSE,
  tmp.img.dir = "/tmp")
```

```{r}
dfSummary(test,
  plain.ascii = FALSE, 
  style = "grid", 
  graph.magnif = 0.75,
  valid.col = FALSE,
  tmp.img.dir = "/tmp")
```

```{r}
test %>% anti_join(train %>% select(Producto_ID) %>% distinct()) %>% count()
```
Searching for new products in the test set, there are 20 new products which is a really small portion of the total 1.3M of test products.

### Wrangling

```{r}
clients %<>% 
  group_by(Cliente_ID) %>% 
  summarise(NombreCliente = first(NombreCliente))
```
Since Data Files set is properly tied to Cliente_ID, for our purposes we can safely discard those repeated names and use the first occurrence of a Cliente_ID - NombreCliente pair as a person.

To capture the greater or equal than 0 effect in demand, we smooth Demanda_uni_equil logarithmically.
We also use an artificiall 0 demand for testing set, to avoid the problems NA may arise.
And finally add the 9 
```{r}
train %<>% mutate(Demanda_uni_equil = log1p(Demanda_uni_equil), Type = "train")
test %<>% mutate(Demanda_uni_equil = 0, Type = "test")

DF <- train %>% filter(Semana == 8) %>% 
  bind_rows(test)

#now compute mean demand + count on train weeks 3 - 7 then join with weeks 8, 9:
DF %<>% left_join((train %>% 
    filter(Semana < 8) %>% 
    group_by(Producto_ID, Cliente_ID) %>% 
    summarise(
      m_PC = mean(Demanda_uni_equil),
      N_PC= n())), by = c("Producto_ID", "Cliente_ID")) %>% 
  left_join((train %>% 
    filter(Semana < 8) %>% 
    group_by(Producto_ID) %>% 
    summarise(
      m_P = mean(Demanda_uni_equil),
      N_P = n())), by = c("Producto_ID")) %>% 
  left_join((train %>% 
    filter(Semana < 8) %>% 
    group_by(Cliente_ID) %>% 
    summarise(
      m_C = mean(Demanda_uni_equil),
      N_C= n())), by = c("Cliente_ID"))
```


### Splitting

Since data is already splitted we can skip this step.

## Model Selection

Using an educated guest, the proposed algorithm is an eXtreme Gradient BOOSTing, noting the following:
  - While we do have a sequential prediction task, our data could be characterised as numerical big data.
  - 8 time observations makes it a time-series, GLM or regression based approach unfeasible.
  - Aside from weeks, we only have categorical covariables which aren't suited to give ordinal predictability.
  - As we do have several observations for a given week, we can aggregate summary statistics grouping by covariables.

```{r}
library(xgboost)
```

## Training

```{r}

#boosting on week 9 of train - around 10 million training examples:
D_label <- DF %>% 
  filter(Type == "train") %>% 
  select(Demanda_uni_equil) %>% unlist()

D_ <- as.matrix(DF %>% 
    filter(Type == "train") %>% 
    select(-Demanda_uni_equil, -ID, -Type))
D_data <- xgb.DMatrix(D_,
  label = D_label)

model <- xgb.train(params=list(  
    objective="reg:linear", 
    booster = "gbtree",
    eta=0.5, # 0.1, 
    max_depth=8, # 10, 
    subsample=0.85,
    colsample_bytree=0.7) ,
  data = D_data, 
  nrounds = 25, # 75, 
  verbose = 1,
  print_every_n=5,
  maximize=FALSE,
  eval_metric='rmse')

```

## Evaluation
  - SHAP
  - RMSE
  - xgb.plot
  

```{r}
require("Ckmeans.1d.dp")

importance <- xgb.importance(
  feature_names = colnames(D_data),
  model = model
) 
xgb.ggplot.importance(importance)
```

```{r}
xgb.ggplot.deepness(model)
```

## Prediction
```{r}
D_pred <- predict(model, as.matrix(DF %>% filter(Type == "test") %>% 
    select(-Demanda_uni_equil, -ID, -Type)))
D_pred = expm1(D_pred)
D_pred[D_pred < 0] = 0
DF_hat <- DF %>% filter(Type == "test") %>% 
  select(ID, Semana, Agencia_ID, Cliente_ID, Producto_ID) %>% 
  mutate(Demanda_uni_equil = D_pred)

DF_ <- train %>% filter(Type == "train") %>% 
  select(ID, Semana, Agencia_ID, Cliente_ID, Producto_ID, Demanda_uni_equil) %>% 
  bind_rows(DF_hat)
  

write_csv(DF_, "../data/df_[final]_small.csv")
```

## Visualization

Run the shiny app.